Microxanox - an R package for simulating an \(MICR\)obial ecosystem that can occupy \(OX\)ic or \(ANOX\)ic states.

### 2022-04-11 inst/manuscript/manuscript_microxanox.tex
Use of elsevier article class and beginning of adaptation to SoftwareX Article Structure

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Microxanox - an R package to simulate \ldots{}},
  pdfkeywords={metadata quality; data curation; archival; long term storage; R package;},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Microxanox - an R package to simulate \ldots{}}
\author{true \and true}
\date{2022-03-21 14:42:18}

\begin{document}
\maketitle
\begin{abstract}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  set the context for and purpose of the work;
\item
  indicate the approach and methods;
\item
  outline the main results;
\item
  identify the conclusions and the wider implications.
\end{enumerate}
\end{abstract}

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{2}
\tableofcontents
}
\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\{\textgreater\textgreater{} Biological justification and background - one paragraph \textless\textless\}

@@

\{\textgreater\textgreater{} Add a more detailed description of the model and the additions \textless\textless\}

The ODEs are solved using the ``General Solver for Ordinary Differential Equations'' of the \texttt{deSolve} package (\protect\hyperlink{ref-Soetaert2010}{Soetaert, Petzoldt, \& Setzer, 2010}).

\hypertarget{availability-and-accessability-of-the-package}{%
\section{Availability and Accessability of the Package}\label{availability-and-accessability-of-the-package}}

@@

The package will be updated based on further research activities. It will be tested continuously for the last recent, current and devel version of R.
\{\textgreater\textgreater If we want that, we have to include some, even rudimentary, GITHUB tests.\textless\textless\}
\{\textgreater\textgreater RMK: Agreed - let's discuss this when we talk about the package paper.\textless\textless\}

@@

\hypertarget{the-r-package}{%
\section{The R Package}\label{the-r-package}}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Standard R help for each function (\texttt{?COMMAND} in R)
\item
  Two vignettes acompanying the package, also available under RUniverse \href{@LINK_NEEDED}{Vignette 1} and \href{@LINK_NEEDED}{Vignette 2}
\item
  This paper as a third vignette, also available under RUniverse \href{@LINK_NEEDED}{Vignette 3}
\end{enumerate}

\hypertarget{the-modelling-framework}{%
\subsection{The modelling framework}\label{the-modelling-framework}}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  all parameter needed to run a simulation or find a stable state are contained in a single parameter object. This object is created by using the functions \texttt{new\_...\_parameter()}, \texttt{new\_initial\_state()} and \texttt{new\_strain\_parameter()}. Which one of the \texttt{new\_...\_parameter()} functions has to be used when, will be discussed in the section \ref{runsim} and in the \href{LINK_NEEDED}{User Guide}.
\item
  The function call \texttt{run\_...(parameter)} will run the simulation using the parameter as defined in the object \texttt{parameter}.
\item
  The return value of the \texttt{run\_...(parameter)} function is identical to the parameter object plus an additionl slot named \texttt{results} which contains the results of the run
\item
  As this return value contains all parameter, it is possible to re-run the simulatuion by simply running \texttt{run\_...(result)}.
\end{enumerate}

@@

The individual simulation (\texttt{run\_simulation()} function) is the working horse in this package. In this function, the ODEs are solved. The function needs only one parameter - an object as created by the function \texttt{new\_runsim\_parameter()}. One parameter of this object is the \texttt{strain\_parameter} which can be created by the function \texttt{new\_strain\_parameter()}. For a detailed description of the parameter and how they are created please see the User Guide and which acomnpanies the package or is available at \href{@LINK_NEEDED}{User Guide}

After the paramter object has been defined, it can be used in the \texttt{run\_simulation()} function. The function returns an object which is identical to the parameter, except of an additional slot containing the results. This design produces a fully reproducible object as it can be used instead of a parameter object to be fed back into the \texttt{run\_simulation()} function to run the simulation again from the parameter used to generate the results from.

The function \texttt{plot\_dynamics()} plots a single simulation run, as returned from the \texttt{run\_simulation()} function. This functio is only provided as a convenience function to provide a way to easily see the results of a simulation run. An example plot resulting from this function is shown in \ref{fig:plot-dynamics}.

\textbackslash begin\{figure\}

\{\centering \includegraphics[width=500px]{figures/fig1_plot_dynamics}

\}

\textbackslash caption\{Plot of results of a simulation run using the function \texttt{plot\_dynamics()}\$`. Details can be found in the ``User Guide'' section ``Three strains per functional group''.\}\label{fig:plot-dynamics}
\textbackslash end\{figure\}

\hypertarget{find-a-steady-state-of-the-model}{%
\subsubsection{Find a Steady State of the model}\label{find-a-steady-state-of-the-model}}

@@

\hypertarget{extract-stability-measures}{%
\subsection{Extract Stability Measures}\label{extract-stability-measures}}

@@

The first two use cases come from the User Guide and the Partial Reproduction Vignettes. The third comes from a research article that relied on the package.

@@

\{\textgreater\textgreater Take from the user guide.\textless\textless\}

@@

\{\textgreater\textgreater Take from the partial reproduction.\textless\textless\}

@@

This model (as part of the package) has been used in the production of the paper (\protect\hyperlink{ref-REF_NEEDED}{REF NEEDED, 2222}). It can be used for similar studies in this study systems.

\{\textgreater\textgreater More details from the paper.\textless\textless\}

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Bush2017}{}}%
Bush, T., Diao, M., Allen, R. J., Sinnige, R., Muyzer, G., \& Huisman, J. (2017). Oxic-anoxic regime shifts mediated by feedbacks between biogeochemical processes and microbial community dynamics. \emph{Nature Communications}, \emph{8}(1), 789. doi:\href{https://doi.org/10.1038/s41467-017-00912-x}{10.1038/s41467-017-00912-x}

>>>>

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{SoftwareX} % Sets Journal name

\usepackage{lineno} % add

\usepackage{graphicx}
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Microxanox - an R package for simulating an MICRobial ecosystem that can occupy OXic or ANOXic states.},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}

% Pandoc syntax highlighting
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% From pandoc table feature
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

%% \documentclass[preprint,12pt, a4paper]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
\usepackage{lineno}

\usepackage{float}
\restylefloat{table}

\journal{SoftwareX}

%\usepackage{lineno}
%\linenumbers
%\usepackage{setspace}
%\doublespacing
%\usepackage{fancyhdr}
%\fancyhead{}
%\fancyfoot{}
%\fancyhead[CO,CE]{Microxanox R package}
%\pagestyle{fancy}
%

\begin{document}

\begin{frontmatter}

  \title{Microxanox - an R package for simulating an \(MICR\)obial
ecosystem that can occupy \(OX\)ic or \(ANOX\)ic states.}
    \author[University of Zürich]{Rainer M Krug\corref{Corresponding
Author}}
   \ead{Rainer.Krug@uzh.ch \& Rainer@krugs.de} 
    \author[University of Zürich]{Owen L. Petchey}
   \ead{Owen.Petchey@ieu.uzh.ch} 
      \address[University of Zürich]{Department of Evolutionary Biology
and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich}
      \cortext[1]{Corresponding Author}
    \cortext[2]{Equal contribution}

  \begin{abstract}
  Ca. 100 words.
  \end{abstract}
   \begin{keyword} metadata quality; data curation; archival; long term
storage; R package;\end{keyword}
 \end{frontmatter}

\hypertarget{required-metadata}{%
\section{Required Metadata}\label{required-metadata}}

\hypertarget{current-code-version}{%
\subsection{Current code version}\label{current-code-version}}

Ancillary data table required for subversion of the codebase.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{Nr.}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Code metadata description}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Please fill in this column}
\end{minipage} \\
\midrule
\endhead
C1 & Current code version & For example v42 \\
C2 & Permanent link to code/repository used for this code version &
\url{https://github.com/UZH-PEG/microxanox} \\
C3 & Code Ocean compute capsule & \\
C4 & Legal Code License & CC BY 4.0 \\
C5 & Code versioning system used & git \\
C6 & Software code languages, tools, and services used & R \\
C7 & Compilation requirements, operating environments & R (\textgreater=
4.1.0), ADD PACKAGES \\
C8 & If available Link to developer documentation/manual & \\
C9 & Support email for questions & Rainer.Krug@uzh.ch;
Rainer@krugs.de \\
\bottomrule
\end{longtable}

\hypertarget{motivation-and-significance}{%
\section{Motivation and
significance}\label{motivation-and-significance}}

Introduce the scientific background and the motivation for developing
the software. Explain why the software is important, and describe the
exact (scientific) problem(s) it solves. Indicate in what way the
software has contributed (or how it will con- tribute in the future) to
the process of scientific discovery; if available, this is to be
supported by citing a research paper using the software. Provide a
description of the experimental setting (how does the user use the
software?). Introduce related work in literature (cite or list
algorithms used, other software etc.).

\hypertarget{software-description}{%
\section{Software description}\label{software-description}}

Describe the software in as much as is necessary to establish a
vocabulary needed to explain its impact.

\hypertarget{software-architecture}{%
\subsection{Software Architecture}\label{software-architecture}}

Give a short overview of the overall software architecture; provide a
pic- torial component overview or similar (if possible). If necessary
provide im- plementation details.

\hypertarget{software-functionalities}{%
\subsection{Software Functionalities}\label{software-functionalities}}

Present the major functionalities of the software.

\hypertarget{illustrative-examples}{%
\section{Illustrative Examples}\label{illustrative-examples}}

Provide at least one illustrative example to demonstrate the major
functions. Optional: you may include one explanatory video that will
appear next to your article, in the right hand side panel.

\hypertarget{impact}{%
\section{Impact}\label{impact}}

This is the main section of the article and the reviewers weight the
description here appropriately Indicate in what way new research
questions can be pursued as a result of the software (if any). Indicate
in what way, and to what extent, the pursuit of existing research
questions is improved (if so). Indicate in what way the software has
changed the daily practice of its users (if so). Indicate how widespread
the use of the software is within and outside the intended user group.
Indicate in what way the software is used in commercial settings and/or
how it led to the creation of spin-off companies (if so).

Set out the conclusion of this original software publication.

\hypertarget{conflict-of-interest}{%
\section{Conflict of Interest}\label{conflict-of-interest}}

We wish to confirm that there are no known conflicts of interest
associated with this publication and there has been no significant
financial support for this work that could have influenced its outcome.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

Optionally thank people and institutes you need to acknowledge.

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

{\{\textgreater\textgreater Biological justification and background -
one paragraph - DONE\textless\textless\} }

Many ecosystems are exposed to gradual changes of environmental
variables, to which the responses are not always as gradual as the
change of the environmental variable. One example where the gradual
change of a single environmental causing an abrupt change of the system
is the switch from an aerobic to anaerobic system. This system hass been
investigated by Bush et al. (\protect\hyperlink{ref-Bush2017}{2017}). We
wanted to take this investigation one step further, and look at the role
of an increased biodiversity plays in these dynamics
(\protect\hyperlink{ref-REF_NEEDED}{REF NEEDED, 2222}). For this
purpose, we developed this package.

@@

{\{\textgreater\textgreater{} Add a more detailed description of the
model and the additions \textless\textless\}
\{\textgreater\textgreater TODO\textless\textless\}}

The ODEs are solved using the ``General Solver for Ordinary Differential
Equations'' of the \texttt{deSolve} package
(\protect\hyperlink{ref-Soetaert2010}{Soetaert et al., 2010}).

\hypertarget{availability-and-accessability-of-the-package}{%
\section{Availability and Accessability of the
package}\label{availability-and-accessability-of-the-package}}

@@

The package will be updated based on further research activities. It
will be tested continuously for the last recent, current and devel
version of R. {\{\textgreater\textgreater If we want that, we have to
include some, even rudimentary, GITHUB tests.\textless\textless\}
\{\textgreater\textgreater RMK: Agreed - let's discuss this when we talk
about the package paper.\textless\textless\} }

@@

\hypertarget{package-description}{%
\section{Package description}\label{package-description}}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Standard R help for each function (\texttt{?COMMAND} in R)
\item
  Two vignettes acompanying the package:
\end{enumerate}

\begin{itemize}
\tightlist
\item
  A \textbf{User Guide}, also available on the R-Universe
  \href{@LINK_NEEDED}{Vignette 1}. This gives an introduction of the
  package and its functions and provides a walk-through of the process
  of running, visualizing and extra and \href{@LINK_NEEDED}{Vignette 2}
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  This paper as a third vignette, also available under RUniverse
  \href{@LINK_NEEDED}{Vignette 3}
\end{enumerate}

\hypertarget{reproducibility}{%
\subsection{Reproducibility}\label{reproducibility}}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  all parameter needed to run a simulation or find a stable state are
  contained in a single parameter object. This object is created by
  using the functions \texttt{new\_...\_parameter()},
  \texttt{new\_initial\_state()} and \texttt{new\_strain\_parameter()}.
  Which one of the \texttt{new\_...\_parameter()} functions has to be
  used when, will be discussed in the section @ref(runsim) and in the
  \href{LINK_NEEDED}{User Guide}.
\item
  The function call \texttt{run\_...(parameter)} will run the simulation
  using the parameter as defined in the object \texttt{parameter}.
\item
  The return value of the \texttt{run\_...(parameter)} function is
  identical to the parameter object plus an additionl slot named
  \texttt{results} which contains the results of the run
\item
  As this return value contains all parameter, it is possible to re-run
  the simulatuion by simply running \texttt{run\_...(result)}.
\end{enumerate}

@@

The individual simulation (\texttt{run\_simulation()} function) is the
working horse in this package. In this function, the ODEs are solved.
The function needs only one argument - an object as created by the
function \texttt{new\_runsim\_parameter()}. One parameter of this object
is the \texttt{strain\_parameter} which can be created by the function
\texttt{new\_strain\_parameter()}. For a detailed description of the
parameter and how they are created please see the User Guide and which
acomnpanies the package or is available at \href{@LINK_NEEDED}{User
Guide}

After the parameter object has been defined, it can be used in the
\texttt{run\_simulation()} function. The function returns an object
which is identical to the parameter, except of an additional slot
containing the results. This design produces a fully reproducible object
as it can be used instead of a parameter object to be fed back into the
\texttt{run\_simulation()} function to run the simulation again from the
parameter used to generate the results from.

The function \texttt{plot\_dynamics()} plots a single simulation run, as
returned from the \texttt{run\_simulation()} function. This functio is
only provided as a convenience function to provide a way to easily see
the results of a simulation run. An example plot resulting from this
function is shown in @ref(fig:plot-dynamics).

\begin{figure}

{\centering \includegraphics[width=500px]{figures/ug_three_strains_dynamics}

}

\caption{Plot of results of a simulation run using the function $plot_dynamics()$. Details can be found in the "User Guide" section "Three strains per functional group".}\label{fig:plot-dynamics}
\end{figure}

\hypertarget{finding-a-steady-state-of-the-model}{%
\subsubsection{Finding a Steady State of the
model}\label{finding-a-steady-state-of-the-model}}

@@

\hypertarget{extracting-stability-measures}{%
\subsection{Extracting Stability
Measures}\label{extracting-stability-measures}}

@@

The first two use cases are described in detail in the User Guide and
the Partial Reproduction Vignettes. The third is taken from the REF
NEEDED (\protect\hyperlink{ref-REF_NEEDED}{2222}) for which this R
package was designed. All of these use cases can be expanded to larger
numbers of strains per functional group and variable values van be
changed.

@@

In the \href{LINK_NEEDED}{User Guide} we used a one strain system
(section ``1 strain per functional group'') and three strain system
(section ``3 strains per functional group'') to determine as an example
the stable states during temporal environmental changes (the oxygen
diffusivity). From these simulatiuons, we extracted measures of
nonlinearity and hysteresis. See Fig @ref(fig:plot-dynamics) as an
example plot of the simulations.

\begin{figure}

{\centering \includegraphics[width=500px]{figures/ug_three_strains_stable_state}

}

\caption{Plot of the stable states of the simulation runs under different oxygen diffusivity. The top graph are the Organisms (ech initially with three strains) while the lower graph is the substrate availability under the same oxygen diffusivities. Details can be found in the "User Guide" section "Three strains per functional group".}\label{fig:uc1_stable_state}
\end{figure}

@@

One of the reasons to develop this package was to reproduce the results
presented in Bush et al. (\protect\hyperlink{ref-Bush2017}{2017}), this
was achieved as demonstrated in the \href{LINK_NEEDED}{Partial
Reproduction supplement}. All aspects in the paper could be reproduced
and are shown in the vignette.

\begin{figure}

{\centering \includegraphics[width=500px]{figures/user_guide_hysteresis}

}

\caption{Hysteresis of all assessed combinations of variability.}\label{fig:user_guide_hysteresis}
\end{figure}

@@

As discussed in the paper (\protect\hyperlink{ref-REF_NEEDED}{REF
NEEDED, 2222}), the role biodiversity plays in abrupt regime shifts
based on gradual changing environmental parameter is not well
understood. This model (as part of the package) has been used to
investigate these dynamics and the results are available in REF NEEDED
(\protect\hyperlink{ref-REF_NEEDED}{2222}).

{\{\textgreater\textgreater More details from the
paper.\textless\textless\}\\
\{\textgreater\textgreater Romana will provide two or three
sentences\textless\textless\}}

\hypertarget{conclusions-1}{%
\section{Conclusions}\label{conclusions-1}}

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Bush2017}{}}%
Bush, T., Diao, M., Allen, R.J., Sinnige, R., Muyzer, G., Huisman, J.,
2017. Oxic-anoxic regime shifts mediated by feedbacks between
biogeochemical processes and microbial community dynamics. Nature
Communications 8, 789.
doi:\href{https://doi.org/10.1038/s41467-017-00912-x}{10.1038/s41467-017-00912-x}

\leavevmode\vadjust pre{\hypertarget{ref-REF_NEEDED}{}}%
REF NEEDED, A., 2222. {REFERNECE NEEDED}. Journal of missing references.
